\bf Visualising How Non-linear Dimension Reduction Warps Your Data

### 2024-01-14 paper.qmd
Add links to the model of S-curve, and What I lear from the models

```{r}
#| warning: false
#| echo: false
#| label: fig-modelScurve
#| fig-cap: Model.

@@

$$
\text{MSE} = \sum_{j = 1}^{n} \frac{\sum_{i = 1}^{p} (x_{ij} - C_{x_ij})^2}{\text{total number of bins}}
$$ {#eq-equation9}

@@

>>>>

@@

$$
\text{MSE} = \sum_{j = 1}^{n} \frac{\sum_{i = 1}^{p} (x_{ij} - C_{x_ij})^2}{b}
$$ {#eq-equation9}

@@

### Model interpretation

The visualizations of NLDR techniques for the S-curve dataset shows different representations in 2D, as showcased in Figure @fig-nldervis. The question at hand is determining the best NLDR technique to represent the S-curve data. Examining the 2D layouts in @fig-nldervis suggests that TriMAP or PaCMAP may be best representations. However, the challenge is to decide the best representation with considerable evidence.

After considering the models constructed from the algorithm for each NLDR techniques and visualizing them in high-D space with the original data, shows that all the NLDR techniques capture the non-linearity of the S-curve (see videos of the models linked in @fig-modelScurve). But thorough examination of each model in high-D space shows some surprising findings. The model of PHATE is flattened and lack of capturing the width of S-curve (see videos of the model with PHATE linked in @fig-modelScurve). This nature of PHATE is also evicted by the considerably higher AIC and MSE values for PHATE as shown in @fig-diagnosticpltScurve. Furthermore, there are two edges that are not connected to any points in the model of TriMAP (see see videos of the model with TriMAP linked in @fig-modelScurve). This model shows how the patterns in 2D representation of TriMAP capture by the model (see @fig-nldervis (d)). As shown in @fig-nldervis (b), UMAP have four clusters in a curvilinear structure. However, the model reveals that even though UMAP shows a curvilinear in 2D, UMAP actually capture the structure of the S-curve without any disjoint in the model (see see videos of the model with UMAP linked in @fig-modelScurve). In the 2D layouts of tSNE and PaCMAP (see @fig-nldervis (a), (e)), there are some sparse areas that are captured by the models effectively (see see videos of the model with tSNE and PaCMAP linked in @fig-modelScurve).

Different NLDR techniques perform differently on the S-curve data and tSNE provides the best representation in 2D according to AIC and MSE values (see @fig-diagnosticpltScurve).

```{r}
#| warning: false
#| echo: false
#| label: fig-modelScurve
#| fig-cap: Is there a best model to represent the original data in 2D space or are they all providing equivalent information?, (a) Model in the 2D space with tSNE (<https://youtu.be/uy9SkAo6gAE>), (b) Model in the 2D space with UMAP (<https://youtu.be/0MJDhHrh_Ug>), (c) Model in the 2D space with PHATE (<https://youtu.be/HbVv0uy0QWk>), (d) Model in the 2D space with TriMAP (<https://youtu.be/2OGwipAjzc8>), and (e) Model in the 2D space with PaCMAP (<https://youtu.be/pkokI8d-cBk>).

### 2024-01-14 paper.tex
Add links to the model of S-curve, and What I lear from the models

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, interior hidden, breakable, enhanced, sharp corners, frame hidden, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi

@@

\caption{\label{fig-modelScurve}Model.}

@@

\begin{equation}\protect\hypertarget{eq-equation9}{}{
\text{MSE} = \sum_{j = 1}^{n} \frac{\sum_{i = 1}^{p} (x_{ij} - C_{x_ij})^2}{\text{total number of bins}}
}\label{eq-equation9}\end{equation}

@@

>>>>

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, interior hidden, boxrule=0pt, sharp corners, frame hidden, breakable, enhanced]}{\end{tcolorbox}}\fi

@@

@@

\begin{equation}\protect\hypertarget{eq-equation9}{}{
\text{MSE} = \sum_{j = 1}^{n} \frac{\sum_{i = 1}^{p} (x_{ij} - C_{x_ij})^2}{b}
}\label{eq-equation9}\end{equation}

@@

\hypertarget{model-interpretation}{%
\subsection{Model interpretation}\label{model-interpretation}}

The visualizations of NLDR techniques for the S-curve dataset shows
different representations in 2D, as showcased in Figure
Figure~\ref{fig-nldervis}. The question at hand is determining the best
NLDR technique to represent the S-curve data. Examining the 2D layouts
in Figure~\ref{fig-nldervis} suggests that TriMAP or PaCMAP may be best
representations. However, the challenge is to decide the best
representation with considerable evidence.

After considering the models constructed from the algorithm for each
NLDR techniques and visualizing them in high-D space with the original
data, shows that all the NLDR techniques capture the non-linearity of
the S-curve (see videos of the models linked in
Figure~\ref{fig-modelScurve}). But thorough examination of each model in
high-D space shows some surprising findings. The model of PHATE is
flattened and lack of capturing the width of S-curve (see videos of the
model with PHATE linked in Figure~\ref{fig-modelScurve}). This nature of
PHATE is also evicted by the considerably higher AIC and MSE values for
PHATE as shown in Figure~\ref{fig-diagnosticpltScurve}. Furthermore,
there are two edges that are not connected to any points in the model of
TriMAP (see see videos of the model with TriMAP linked in
Figure~\ref{fig-modelScurve}). This model shows how the patterns in 2D
representation of TriMAP capture by the model (see
Figure~\ref{fig-nldervis} (d)). As shown in Figure~\ref{fig-nldervis}
(b), UMAP have four clusters in a curvilinear structure. However, the
model reveals that even though UMAP shows a curvilinear in 2D, UMAP
actually capture the structure of the S-curve without any disjoint in
the model (see see videos of the model with UMAP linked in
Figure~\ref{fig-modelScurve}). In the 2D layouts of tSNE and PaCMAP (see
Figure~\ref{fig-nldervis} (a), (e)), there are some sparse areas that
are captured by the models effectively (see see videos of the model with
tSNE and PaCMAP linked in Figure~\ref{fig-modelScurve}).

Different NLDR techniques perform differently on the S-curve data and
tSNE provides the best representation in 2D according to AIC and MSE
values (see Figure~\ref{fig-diagnosticpltScurve}).

\caption{\label{fig-modelScurve}Is there a best model to represent the
original data in 2D space or are they all providing equivalent
information?, (a) Model in the 2D space with tSNE
(\url{https://youtu.be/uy9SkAo6gAE}), (b) Model in the 2D space with
UMAP (\url{https://youtu.be/0MJDhHrh_Ug}), (c) Model in the 2D space
with PHATE (\url{https://youtu.be/HbVv0uy0QWk}), (d) Model in the 2D
space with TriMAP (\url{https://youtu.be/2OGwipAjzc8}), and (e) Model in
the 2D space with PaCMAP (\url{https://youtu.be/pkokI8d-cBk}).}
